Supporting online material for\\ \emph{Effectiveness of UNAIDS targets and HIV vaccination across 127 countries}

### 2016-08-21 Documents/supplementary_text.tex
More work on supplement.

\usepackage[font = {sf}, labelfont = {bf},
  labelsep = period, singlelinecheck = off]{caption}
\usepackage[font = {sf}]{floatrow}
\usepackage{tabularx}

@@

\usepackage[style = science, article-title = true, maxnames = 100]{biblatex}
\addbibresource{supplementary_text.bib}
\newcommand{\labeltitlepunc}{}
\DeclareBibliographyDriver{incollection}{%
  \usebibmacro{bibindex}%
  \usebibmacro{begentry}%
  \usebibmacro{author/translator+others}%
  \setunit{\labelnamepunct}\newblock
  \printlist{language}%
  \newunit\newblock
  \usebibmacro{byauthor}%
  \newunit\newblock
  \usebibmacro{title}%
  \usebibmacro{in:}%
  \usebibmacro{maintitle+booktitle}%
  \newunit\newblock
  \usebibmacro{byeditor+others}%
  \newunit\newblock
  \printfield{volumes}%
  \newunit\newblock
  \printfield{note}%
  \newunit\newblock
  \usebibmacro{publisher+location+date}%
  \newunit\newblock
  \printfield{volume}
  \newunit
  \usebibmacro{chapter+pages}%
  \newunit\newblock
  \iftoggle{bbx:isbn}
    {\printfield{isbn}}
    {}%
  \newunit\newblock
  \usebibmacro{doi+eprint+url}%
  \newunit\newblock
  \usebibmacro{addendum+pubstate}%
  \setunit{\bibpagerefpunct}\newblock
  \usebibmacro{pageref}%
  \newunit\newblock
  \usebibmacro{related}%
  \usebibmacro{finentry}%
}

@@

@@

We developed a continuous-time, compartmental model that divides
people aged $15$\;y and older into non-overlapping HIV states.  We
chose to consider only these ages because they are the vast majority
of PLHIV, 93\% in 2014 \cite{UNICEF}, and because HIV and demographic
data are often stratified into these ages, although some data used to
parametrize the model was only available for ages $15$--$49$ rather
than for ages $15$ and older.  The background mortality rate ($\mu$)
in each of the 127 countries was collected from published estimates
\cite{World_Development_Indicators2013-ee}.  The recruitment rate into
ages $15\;\text{y}$ and older group ($\kappa$) was set by the difference
between the country's population growth rate \cite{WorldBankpg} and
its background mortality rate
\cite{World_Development_Indicators2013-ee}, as would be true if the
population was at its stable age structure.

@@

We modeled 3 different targets for the proportion diagnosed ($p_D$), the
proportion treated ($p_T$), and the proportion with viral suppression ($p_V$).
\begin{description}
\item[Status quo] Given initial proportions $p_D(2015)$, $p_T(2015)$,
  and $p_V(2015)$, the targets are for the proportion to remain fixed
  at that level for all times:
  \begin{equation}
    \label{status_quo_target}
    \begin{split}
      p_D^*(t) &= p_D(2015), \\
      p_T^*(t) &= p_T(2015), \\
      p_V^*(t) &= p_V(2015).
    \end{split}
  \end{equation}

@@

For vaccination, we assume that vaccination rollout begins in year
$t_V = 2020$ (or $t_V = 2025$ in vaccine scenario analysis,
\autoref{uncertainty}), that rollout linearly increases the proportion
of non-infected population covered from 0\% by $r_V = 25\%$ per year
(or $r_V = 10\%$ per year) up to $c_V = 70\%$ (or $c_V = 50\%$ or
$c_V = 90\%$) and staying constant thereafter.  Thus, the vaccination
target is
\begin{equation}
  \label{vaccination_target}
  p_R^*(t) =
  \begin{cases}
    0 & \text{if $t < t_V$},
    \\
    r_V (t - t_V) & \text{if $t_V \leq t < T_V$},
    \\
    c_V & \text{if $t \geq T_V$},
  \end{cases}
\end{equation}
where $T_V = t_V + c_V / r_V$ is the time when rollout reaches
the full coverage $c_V$.

Projections were generated for each country through model simulations
using 1000 random samples from published parameter distributions
(table S1) and the country-specific estimated distribution of the
transmission rate. We evaluated the global and country-level impacts
of meeting the UNAIDS targets and vaccine deployment on four outcome
measures: HIV incidence, PLHIV, PLAIDS, and AIDS-related deaths
between 2015 and 2035. Our base-case scenario assumed maintenance of
the status quo: that 2015 levels of the proportion of PLHIV diagnosed
(pD), proportion of those diagnosed on ART (pT), and proportion of
those on treatment who have achieved viral suppression (pV) continue
unchanged until 2035. The intervention scenarios were (1.) the
attainment of the 90–90–90 target by linear increases in pD, pT, and
pV until 2020 followed by maintenance of those levels until 2035, and
(2.) linear increase in the diagnosis and treatment levels to the
90–90–90 target by 2020, a linear increase from 90–90–90 to the
95–95–95 target by 2030, and maintenance of the 95–95–95 levels to
2035. (For 90–90–90, if any initial level was above 90\%, it was kept
constant going forward so that the intervention did not worsen
coverage.  Likewise, for 95–95–95, any initial level above 95\% was
kept constant.)  For vaccination, we assumed deployment of a vaccine
with 50\% efficacy in 2020 with coverage increasing linearly by 25\%
per year, continuing up to 70\% at 2.8 years post-rollout, and
maintaining constant 70\% coverage thereafter.  We also performed a
partial rank correlation coefficient (PRCC) sensitivity analysis for
each model parameter and a scenario analysis to account for
uncertainty in the profile of the prospective vaccine.

@@

For 127 countries, we found sufficient data to parametrize our model:
these data consisted of estimates of longitudinal HIV incidence and
prevalence for ages 15–49 years starting as early as 1990 in some
countries and running through 2015, along with basic demographics.

Several sources provided information to parameterize our model at the
country level. Demographic data, including population growth rate
\cite{WorldBankpg}, death rate
\cite{World_Development_Indicators2013-ee}, and number of people aged
15--49 years \cite{The_World_Bank2016-fd} were obtained from the World
Bank. Longitudinal HIV incidence and prevalence estimates for the
15--49 age group, spanning from as early as 1990 to 2015, was
primarily derived from the AIDSinfo database produced by UNAIDS
\cite{Unaids2016-an}. Other sources included UNAIDS Country Progress
Reports \cite{Unaids2016-am} published from 2012 to 2016 and AIDS Data
Hub \cite{AIDSdatahub-fg}, which compiles data from UNAIDS, UNICEF,
WHO, and the Asian Development Bank. These sources were also used to
inform the initial conditions for the model: the number of people who
have been diagnosed with HIV, who are on ART, and who have viral
suppression or have been retained on treatment for at least 12
months. Where data were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  We found sufficient data to
parametrize our mathematical modeling for 127 countries.  See
Supplemental data files for full information on data sources for each
country.  The data files are also available in the public source-code
repository \cite{medlock2016-git}.

@@

\section{Source code}

The model simulation and analysis tools, written in Python, are
publicly available at
\url{https://github.com/janmedlock/HIV-95-vaccine/}
\cite{medlock2016-git}.

>>>>

\usepackage[font={sf}, labelfont={bf},
  labelsep=period, singlelinecheck=off]{caption}
\usepackage[font={sf}]{floatrow}
\usepackage{tabularx}

@@

\usepackage[style=science, article-title=true, maxnames=100]{biblatex}
\addbibresource{supplementary_text.bib}
\newcommand{\labeltitlepunc}{}
\DeclareBibliographyDriver{incollection}{%
  \usebibmacro{bibindex}%
  \usebibmacro{begentry}%
  \usebibmacro{author/translator+others}%
  \setunit{\labelnamepunct}\newblock
  \printlist{language}%
  \newunit\newblock
  \usebibmacro{byauthor}%
  \newunit\newblock
  \usebibmacro{title}%
  \usebibmacro{in:}%
  \usebibmacro{maintitle+booktitle}%
  \newunit\newblock
  \usebibmacro{byeditor+others}%
  \newunit\newblock
  \printfield{volumes}%
  \newunit\newblock
  \printfield{note}%
  \newunit\newblock
  \usebibmacro{publisher+location+date}%
  \newunit\newblock
  \printfield{volume}
  \newunit
  \usebibmacro{chapter+pages}%
  \newunit\newblock
  \iftoggle{bbx:isbn}
    {\printfield{isbn}}
    {}%
  \newunit\newblock
  \usebibmacro{doi+eprint+url}%
  \newunit\newblock
  \usebibmacro{addendum+pubstate}%
  \setunit{\bibpagerefpunct}\newblock
  \usebibmacro{pageref}%
  \newunit\newblock
  \usebibmacro{related}%
  \usebibmacro{finentry}%
}

@@

The model simulation and analysis tools, written in Python, are
publicly available at\\
\url{https://github.com/janmedlock/HIV-95-vaccine/}
\cite{medlock2016-git}.

@@

We developed a continuous-time, compartmental model that divides
people aged $15$\;y and older into non-overlapping HIV states.  We
chose to consider only these ages because they are the vast majority
of PLHIV---93\% in 2014 \cite{UNICEF}---and because HIV and
demographic data are often stratified into these ages, although some
data used to parametrize the model was only available for ages
$15$--$49$ rather than for ages $15$ and older.  The background
mortality rate ($\mu$) in each of the 127 countries was collected from
published estimates \cite{World_Development_Indicators2013-ee}.  The
recruitment rate ($\kappa$) into the age $15\;\text{y}$ and older
group was set by the difference between the country's population
growth rate \cite{WorldBankpg} and its background mortality rate
\cite{World_Development_Indicators2013-ee}, as would be true if the
population was at its stable age structure.

@@

We modeled 3 different targets for the proportion diagnosed ($p_D$), the
proportion treated ($p_T$), and the proportion with viral suppression ($p_V$).
\begin{description}
\item[Status quo] The status quo target is for the proportions $p_D$,
  $p_T$, and $p_V$ to remain fixed at their initial levels going
  forward:
  \begin{equation}
    \label{status_quo_target}
    \begin{split}
      p_D^*(t) &= p_D(2015), \\
      p_T^*(t) &= p_T(2015), \\
      p_V^*(t) &= p_V(2015).
    \end{split}
  \end{equation}

@@

For vaccination, we assume that rollout begins in year $t_V = 2020$
(or $t_V = 2025$ in vaccine scenario analysis, \autoref{uncertainty}),
that rollout linearly increases the proportion of the non-infected
population covered from 0\% by $r_V = 25\%$ per year (or $r_V = 10\%$
per year) up to $c_V = 70\%$ (or $c_V = 50\%$ or $c_V = 90\%$) and
staying constant thereafter.  Thus, the vaccination target is
\begin{equation}
  \label{vaccination_target}
  p_R^*(t) =
  \begin{cases}
    0 & \text{if $t < t_V$},
    \\
    r_V (t - t_V) & \text{if $t_V \leq t < T_V$},
    \\
    c_V & \text{if $t \geq T_V$},
  \end{cases}
\end{equation}
where $T_V = t_V + c_V / r_V$ is the time when rollout reaches
the ultimate coverage $c_V$.

Target strategies were one of the three diagnosis and treatment
targets (status quo, 90--90--90, or 95--95--95) combined with either
no vaccination ($p_R^*(t) = 0$) or vaccination at some level of
efficacy, ultimate coverage, start date of rollout, and speed of
rollout.

@@

For 127 countries, we found sufficient data to parametrize our model.
Demographic data, including population growth rate \cite{WorldBankpg},
death rate \cite{World_Development_Indicators2013-ee}, and number of
people aged 15--49 years \cite{The_World_Bank2016-fd} were obtained
from the World Bank. Longitudinal HIV incidence and prevalence
estimates for ages 15--49, spanning from as early as 1990 to 2015,
were primarily derived from the AIDSinfo database produced by UNAIDS
\cite{Unaids2016-an}. Other sources included UNAIDS Country Progress
Reports \cite{Unaids2016-am} published from 2012 to 2016 and AIDS Data
Hub \cite{AIDSdatahub-fg}, which compiles data from UNAIDS, UNICEF,
WHO, and the Asian Development Bank. These sources were also used to
inform the initial conditions for the model: the number of people
diagnosed with HIV, the number on ART, and the number who have viral
suppression or have been retained on treatment for at least 12
months. Where data were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  See Tables S2--S6 for full
information on data sources for each country.  The data tables are
also available in the public source-code repository
\cite{medlock2016-git}.

@@

Projections were generated for each country through model simulations
using 1000 random samples from published parameter distributions
(\autoref{model_param}) and the country-specific estimated
distribution of the transmission rate (\autoref{model_fitting}). We
evaluated the global and country-level impacts of meeting the UNAIDS
targets and vaccine deployment on four outcome measures: HIV
incidence, PLHIV, PLAIDS, and AIDS-related deaths between 2015 and
2035. Our base-case scenario assumed maintenance of the status quo:
that 2015 levels of the proportion of PLHIV diagnosed (pD), proportion
of those diagnosed on ART (pT), and proportion of those on treatment
who have achieved viral suppression (pV) continue unchanged until
2035. The intervention scenarios were (1.) the attainment of the
90–90–90 target by linear increases in pD, pT, and pV until 2020
followed by maintenance of those levels until 2035, and (2.) linear
increase in the diagnosis and treatment levels to the 90–90–90 target
by 2020, a linear increase from 90–90–90 to the 95–95–95 target by
2030, and maintenance of the 95–95–95 levels to 2035. (For 90–90–90,
if any initial level was above 90\%, it was kept constant going
forward so that the intervention did not worsen coverage.  Likewise,
for 95–95–95, any initial level above 95\% was kept constant.)  For
vaccination, we assumed deployment of a vaccine with 50\% efficacy in
2020 with coverage increasing linearly by 25\% per year, continuing up
to 70\% at 2.8 years post-rollout, and maintaining constant 70\%
coverage thereafter.  We also performed a partial rank correlation
coefficient (PRCC) sensitivity analysis for each model parameter and a
scenario analysis to account for uncertainty in the profile of the
prospective vaccine.
